---
title: "Genetic interactions CellMap exploration, multicore version"
author: "H Qin"
date: "11/1 - 11/9 /2017 "
output: html_document
---
Multicore 
See http://hongqinlab.blogspot.com/2017/06/yeast-genetic-map.html

CELLMAP provids genetics interaction in pairwise format. 
The interactions are provided for strain_ids (alleles). 
The mapping of strain_id and ORF is in "strain_ids_and_single_mutant_fitness.csv" . 
```{r}
rm(list=ls());
set.seed(2017);
datapath = "~/github/cellmap_parent/CellMap/20170626/S1.pairwise/"; #ridgeside
#datapath = "~/data/Sce/CellMap/20170626/S1.pairwise/"; #applejack
debug = 0;
list.files(path=datapath);

library(foreach)
library(doMC)
registerDoMC(4)
```

Load naming lookup tables
```{r}
dic = read.csv(paste(datapath, "strain_ids_and_single_mutant_fitness.csv", sep=''))
```

Load essential and non-essential infor that H. Qin generated. 
```{r}
list.files(path="data");
fit = read.csv("data/SummaryRegressionHetHomFactorized2015Oct13.csv")
```

Load pairwise interaction data
```{r}
#Essential X Essential 
tb.ee = read.table(paste(datapath,"SGA_ExE.txt", sep=''), header=T, sep="\t");
summary(tb.ee);
```

```{r}
#EXN and NXE
tb.en = read.table(paste(datapath,"SGA_ExN_NxE.txt", sep=''), header=T, sep="\t");
summary(tb.en);
```

```{r}
#NXN I do not need consider NxN for my aging modeling project
tb.nn = read.table(paste(datapath,"SGA_NxN.txt", sep=''), header=T, sep="\t");
summary(tb.nn);
```

Columns names in the 3 tables are the same. 
```{r}
rbind( names(tb.en), names(tb.ee), names(tb.nn))
```

Merge 3 tables into 1 table. 
```{r}
tb.gin = rbind(tb.ee, tb.en, tb.nn);
```
Double-check the merged results
```{r}
length(tb.gin[,1]) == sum(length(tb.ee[,1]), length(tb.en[,1]), length(tb.nn[,1]))
```

Remove unused table to free up memory
```{r}
if (debug==0){
 ls()
 rm(tb.en, tb.ee, tb.nn)
 ls()
}
```


Costanzo2016 suggested lenient, intermediate, and stringennt ways for gin quality check. 

```{r}
tb.gin.lenient = tb.gin[ tb.gin$P.value<=0.05, ];
if (debug ==0) { rm(tb.gin); } #freeup memory
```

Map strain IDs to ORFs, add my essentialFalgs

```{r}
tb.gin.lenient$ORF1 = as.character( dic$Systematic.gene.name[match( tb.gin.lenient$Query.Strain.ID, dic$Strain.ID)] )
tb.gin.lenient$ORF2 = as.character( dic$Systematic.gene.name[match( tb.gin.lenient$Array.Strain.ID, dic$Strain.ID)] )

tb.gin.lenient$essenflag1 = fit$essenflag[ match(tb.gin.lenient$ORF1, fit$orf)]
tb.gin.lenient$essenflag2 = fit$essenflag[ match(tb.gin.lenient$ORF2, fit$orf)]

head(tb.gin.lenient)
```

I need to take only unqiue interactions. This however raise questions how to store score andn p-values. 
I will store the unique key_pairs first. 


```{r}
# i=199990
#i=5347
#i=5348

#for (i in 5347:6999) {
total = length(tb.gin.lenient[,1]);
for (i in 1:total) {
 if( (round(i/1000)*1000 - i)==0 ) { print(paste(i, "::", total - i) ); }  
 pairs = tb.gin.lenient[i , c("ORF1", "ORF2")];
 if ( ! is.na(pairs[1]) ) {
  ordered_pairs = sort(pairs)
  tb.gin.lenient$ordered_pairs[i] =  paste( ordered_pairs[1], ordered_pairs[2], sep="_");
 } else {
  tb.gin.lenient$ordered_pairs[i] = "NA_found"; 
 }
}
```

How many unique interactions (gene pairs) ? 
1394733 in unique_pairs, which is 88.86% of the tb.gin.lenient rows
```{r}
unique_pairs = unique( tb.gin.lenient$ordered_pairs ) 
length(unique_pairs) / length(tb.gin.lenient[,1])
```

Using author's definition to find intermediate and stringent subsets. 
```{r}
tb.gin.intermediate = tb.gin.lenient[ abs(tb.gin.lenient$Genetic.interaction.score..ε.) >0.08, ];
tb.gin.stringent = tb.gin.lenient[ tb.gin.lenient$Genetic.interaction.score..ε.>0.16 | tb.gin.lenient$Genetic.interaction.score..ε.< -0.12, ]
```

```{r}
hist(tb.gin.lenient$Genetic.interaction.score..ε., breaks = 100)
```


```{r}
summary(tb.gin.lenient);
summary(tb.gin.intermediate);
summary(tb.gin.stringent)
```

output
```{r}
write.csv(tb.gin.lenient, "../cellmap_data/gin_lenient20171101.csv", row.names = F);
write.csv(tb.gin.intermediate, "../cellmap_data/gin_intermediate20171101.csv", row.names = F);
write.csv(tb.gin.stringent, "../cellmap_data/gin_stringent20171101.csv", row.names = F);
```



Generate "lenient" unique GINx using average values
```{r lenient GIN}
#Generate unique GINs using match(). This does not average the genetic interaction score
tb.gin.lenient.uniqueByMatch = tb.gin.lenient[match(unique_pairs, tb.gin.lenient$ordered_pairs),  ]
tb.gin.lenient.uniqueByMatch = tb.gin.lenient.uniqueByMatch[ !is.na(tb.gin.lenient.uniqueByMatch$ordered_pairs), ]
tb.gin.lenient.uniqueByMatch = tb.gin.lenient.uniqueByMatch[ !is.na(tb.gin.lenient.uniqueByMatch$Query.Strain.ID), ]
tb.gin.lenient.uniqueByAverage = tb.gin.lenient.uniqueByMatch; 
tb.gin.lenient.uniqueByAverage$buffer = "NA"
total = length(tb.gin.lenient.uniqueByAverage[,1]); #error 20171113
for (i in 1:300) {
 if( (round(i/100)*100 - i)==0 ) { print(paste(i, ":leniet:", total - i) ); }  
 tmp = tb.gin.lenient[ tb.gin.lenient$ordered_pairs == unique_pairs[i],  ]
 if( length(tmp[,1]) > 1) {#more than 1 records, take average
   tb.gin.lenient.uniqueByAverage$buffer[i]  = paste( tmp$Query.Strain.ID, tmp$Array.Strain.ID, sep=" ", collapse="::");
   tb.gin.lenient.uniqueByAverage$Genetic.interaction.score..ε.[i] = mean(tmp$Genetic.interaction.score..ε.)
   tb.gin.lenient.uniqueByAverage$P.value[i] = mean(tmp$P.value)
 }  
 #else do nothing
}
```

Generate "intermediate" unique GINx using average values
```{r intermediate GIN}
tb.gin.intermediate.uniqueByMatch = tb.gin.intermediate[match(unique_pairs, tb.gin.intermediate$ordered_pairs),  ]
tb.gin.intermediate.uniqueByMatch = tb.gin.intermediate.uniqueByMatch[ !is.na(tb.gin.intermediate.uniqueByMatch$ordered_pairs), ]
tb.gin.intermediate.uniqueByMatch = tb.gin.intermediate.uniqueByMatch[ !is.na(tb.gin.intermediate.uniqueByMatch$Query.Strain.ID), ]
tb.gin.intermediate.uniqueByAverage = tb.gin.intermediate.uniqueByMatch; 
tb.gin.intermediate.uniqueByAverage$buffer = "NA"
totalIntermediate = length(tb.gin.intermediate.uniqueByMatch[,1]); #20171113 correction
for (i in 1:500) {
 if( (round(i/100)*100 - i)==0 ) { print(paste(i, ":intermediate:", totalIntermediate - i) ); }  
 tmp = tb.gin.intermediate[ tb.gin.intermediate$ordered_pairs == unique_pairs[i],  ]
 if( length(tmp[,1]) > 1) {#more than 1 records, take average
   tb.gin.intermediate.uniqueByAverage$buffer[i]  = paste( tmp$Query.Strain.ID, tmp$Array.Strain.ID, sep=" ", collapse="::");
   tb.gin.intermediate.uniqueByAverage$Genetic.interaction.score..ε.[i] = mean(tmp$Genetic.interaction.score..ε.)
   tb.gin.intermediate.uniqueByAverage$P.value[i] = mean(tmp$P.value)
 }  
 #else do nothing
}
```


Generate "stringent" unique GINx using average values
```{r stringent GIN}
tb.gin.stringent.uniqueByMatch = tb.gin.stringent[match(unique_pairs, tb.gin.stringent$ordered_pairs),  ]
tb.gin.stringent.uniqueByMatch = tb.gin.stringent.uniqueByMatch[ !is.na(tb.gin.stringent.uniqueByMatch$ordered_pairs), ]
tb.gin.stringent.uniqueByMatch = tb.gin.stringent.uniqueByMatch[ !is.na(tb.gin.stringent.uniqueByMatch$Query.Strain.ID), ]

tb.gin.stringent.uniqueByAverage = tb.gin.stringent.uniqueByMatch; 
tb.gin.stringent.uniqueByAverage$buffer = "NA"
totalstringent = length(tb.gin.stringent.uniqueByMatch[,1]);
for (i in 1:400) {
 if( (round(i/100)*100 - i)==0 ) { print(paste(i, ":stringent:", totalstringent - i) ); }  
 tmp = tb.gin.stringent[ tb.gin.stringent$ordered_pairs == unique_pairs[i],  ]
 tmp = tmp[ !is.na(tmp$Query.Strain.ID), ]
 if( length(tmp[,1]) > 1) {#more than 1 records, take average
   tb.gin.stringent.uniqueByAverage$buffer[i]  = paste( tmp$Query.Strain.ID, tmp$Array.Strain.ID, sep=" ", collapse="::");
   tb.gin.stringent.uniqueByAverage$Genetic.interaction.score..ε.[i] = mean(tmp$Genetic.interaction.score..ε.)
   tb.gin.stringent.uniqueByAverage$P.value[i] = mean(tmp$P.value)
 }  
 #else do nothing
}
```



output
```{r output}
write.csv(tb.gin.lenient.uniqueByAverage, "../cellmap_data/gin_lenient_uniquePairs20171113.csv", row.names = F); 
write.csv(tb.gin.intermediate.uniqueByAverage, "../cellmap_data/gin_intermediate_uniquePairs20171113.csv", row.names = F); 
write.csv(tb.gin.stringent.uniqueByAverage, "../cellmap_data/gin_stringent_uniquePairs20171113.csv", row.names = F); 
```

