---
title: "Genetic interactions CellMap exploration"
author: "H Qin"
date: "10/23 - 20/27 /2017 "
output: html_document
---
See http://hongqinlab.blogspot.com/2017/06/yeast-genetic-map.html

CELLMAP provids genetics interaction in pairwise format. 
The interactions are provided for strain_ids (alleles). 
The mapping of strain_id and ORF is in "strain_ids_and_single_mutant_fitness.csv" . 
```{r}
rm(list=ls());
set.seed(2017);
datapath = "~/github/cellmap_parent/CellMap/20170626/S1.pairwise/"; #ridgeside
#datapath = "~/data/Sce/CellMap/20170626/S1.pairwise/"; #applejack
debug = 1;
list.files(path=datapath);
```

Load naming lookup tables
```{r}
dic = read.csv(paste(datapath, "strain_ids_and_single_mutant_fitness.csv", sep=''))
```

Load essential and non-essential infor that H. Qin generated. 
```{r}
list.files(path="data");
fit = read.csv("data/SummaryRegressionHetHomFactorized2015Oct13.csv")
```

Load pairwise interaction data
```{r}
#Essential X Essential 
tb.ee = read.table(paste(datapath,"SGA_ExE.txt", sep=''), header=T, sep="\t");
summary(tb.ee);
```

```{r}
#EXN and NXE
tb.en = read.table(paste(datapath,"SGA_ExN_NxE.txt", sep=''), header=T, sep="\t");
summary(tb.en);
```

```{r}
#NXN I do not need consider NxN for my aging modeling project
tb.nn = read.table(paste(datapath,"SGA_NxN.txt", sep=''), header=T, sep="\t");
summary(tb.nn);
```

Columns names in the 3 tables are the same. 
```{r}
rbind( names(tb.en), names(tb.ee), names(tb.nn))
```

Merge 3 tables into 1 table. 
```{r}
tb.gin = rbind(tb.ee, tb.en, tb.nn);
```
Double-check the merged results
```{r}
length(tb.gin[,1]) == sum(length(tb.ee[,1]), length(tb.en[,1]), length(tb.nn[,1]))
```

Remove unused table to free up memory
```{r}
if (debug==0){
 ls()
 rm(tb.en, tb.ee, tb.nn)
 ls()
}
```


Costanzo2016 suggested lenient, intermediate, and stringennt ways for gin quality check. 

```{r}
tb.gin.lenient = tb.gin[ tb.gin$P.value<=0.05, ];
if (debug ==0) { rm(tb.gin); } #freeup memory
```

Map strain IDs to ORFs, add my essentialFalgs

```{r}
tb.gin.lenient$ORF1 = as.character( dic$Systematic.gene.name[match( tb.gin.lenient$Query.Strain.ID, dic$Strain.ID)] )
tb.gin.lenient$ORF2 = as.character( dic$Systematic.gene.name[match( tb.gin.lenient$Array.Strain.ID, dic$Strain.ID)] )

tb.gin.lenient$essenflag1 = fit$essenflag[ match(tb.gin.lenient$ORF1, fit$orf)]
tb.gin.lenient$essenflag2 = fit$essenflag[ match(tb.gin.lenient$ORF2, fit$orf)]

head(tb.gin.lenient)
```

I need to take only unqiue interactions. This however raise questions how to store score andn p-values. 
I will store the unique key_pairs first. 


```{r}
# This chunk of codes only runs for subset of tb.gin.lenient, but not 
#pairs = tb.gin.lenient[ c(1:2, 100, 1001, 5000, 9000), c("ORF1", "ORF2")]; #debug
#pairs = tb.gin.lenient[ 1:length(tb.gin.lenient[,1]), c("ORF1", "ORF2")];#DOES NOT RUN correctly
#pairs = tb.gin.lenient[ , c("ORF1", "ORF2")]; #DOES NOT RUN correctly
#pairs = tb.gin.lenient[ 1000:1050, c("ORF1", "ORF2")];
#pairs = data.frame(pairs)
#ordered_pairs = t(apply(pairs, 1, sort)); 
#ordered_pairs = data.frame(ordered_pairs);
#names(ordered_pairs) = c('id1', 'id2');
#ordered_pairs$id1 = as.character( ordered_pairs$id1)
#ordered_pairs$id2 = as.character( ordered_pairs$id2)
#tb.gin.lenient$ordered_pairs =  paste( ordered_pairs$id1, ordered_pairs$id2, sep="_")[5:10]
```

```{r}
# i=199990
#i=5347
#i=5348

#for (i in 5347:6999) {
total = length(tb.gin.lenient[,1]);
for (i in 1:total) {
 if( (round(i/1000)*1000 - i)==0 ) { print(paste(i, "::", total - i) ); }  
 pairs = tb.gin.lenient[i , c("ORF1", "ORF2")];
 if ( ! is.na(pairs[1]) ) {
  ordered_pairs = sort(pairs)
  tb.gin.lenient$ordered_pairs[i] =  paste( ordered_pairs[1], ordered_pairs[2], sep="_");
 } else {
  tb.gin.lenient$ordered_pairs[i] = "NA_found"; 
 }
}
```

How many unique interactions (gene pairs) ? 
```{r}
#unique
```

```{r}
tb.gin.intermediate = tb.gin.lenient[ abs(tb.gin.lenient$Genetic.interaction.score..ε.) >0.08, ];
tb.gin.stringent = tb.gin.lenient[ tb.gin.lenient$Genetic.interaction.score..ε.>0.16 | tb.gin.lenient$Genetic.interaction.score..ε.< -0.12, ]
```


```{r}
hist(tb.gin.lenient$Genetic.interaction.score..ε., breaks = 100)
```


```{r}
summary(tb.gin.lenient);
summary(tb.gin.intermediate);
summary(tb.gin.stringent)
```


output
```{r}
write.csv(tb.gin.lenient, "../cellmap_data/gin_lenient20171101.csv", row.names = F);
write.csv(tb.gin.intermediate, "../cellmap_data/gin_intermediate20171101.csv", row.names = F);
write.csv(tb.gin.stringent, "../cellmap_data/gin_stringent20171101.csv", row.names = F);
```

